---
title: "Pollock Effort"
output:
  html_document: default
---

\n JW pulled this data from AKFIN see "data/AK_Effort_Trawl/Gasper Data/Chinook_Bycatch_Data_Description (2).html"for more specific query info. 

### General methods:

\n First half of sample fraction- How many boat days are observed? 

\n Total Effort Data:
\n We have data on total effort for 1991-2017. Shoreside/CV fish tickets for the GOA only. JW removed the CPs because they don’t always issue fish tickets. They instead track their landings via production reports.

\n Observed data:
\n Jordan counted the number of unique haul_dates for each vessel-sector-trip_target_code-year-nmfsarea-month combo, so we have a summary of all observed hauls, and this is summarised into boat days by taking the . For Shoreside delivers, that is based on your sample fraction (ie still a fraction of effort still unobserved). Note that this still doesn’t identify which deliveries were made to tenders (which of course, are not observed, GS has this data from else where and can match in, only occurs after 2006). Jordan created a trip_target_code2 that is “non-P” when gear is NPT because of 1996 ambiguity, recomends we skip 1996 because there are a bunch of trips (N=327) where trip target code is missing. However, we know that if gear type is “NPT” (non-pelagic trawl), it is not a pollock target. So we will say that if trip_target_code is missing but gear is “NPT” then trip_target_code is not pollock (“Not-P”). This leaves us a trivial number of missing trip_target_code in all years but 1996. Data from 1996 is included in this data set but only by filter out NPT gear, so it likely overestimates effort. 

\n 1991-1995: We do not have direct observed effort infomation here, so instead I used the observer coverage chart in our google drive, to estimate the fraciton of total effort that was observed from 1991-1995. This is VERY likely an overestimate, so we may want to take it out.  
\n 1996-2017: We have information on observed trips. Trip-level data is provided, but we have fishing start and landing dates, so by subtracting those we get an estimate of boat days. Jordan is concerned with this because landed date not necessarily equal to when they came into port nor when they stopped fishing, likely they dont land fish the same day they stopped fishing, likely more than a day, shorter the trip the more liklihood for error. 
  
\n Second half of sample fraction- If there is an observer on board how many fish are the checking for CWT?
\n Observer protocols: Observers sample 100% of catch for Chinook and then check 1/10 chinook for a CWT (Jordan is unsure when this policy started, suggested we just use it for the whole time). Bascially 10% of catch is surveyed for CWT, unless the deliver goes to a tender, than 0% of catch is checked for CWT. 

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message=FALSE, warning = FALSE)
library(tidyverse)
library(here)
library(gmodels)
library(gplots)
library(readr)
col.br= colorRampPalette(c("white", "blue","purple","red")) #colors for heatmap
every_nth = function(n) {
  return(function(x) {x[c(TRUE, rep(FALSE, n - 1))]})
}
```

```{r data prep}
#observed trips 1996- 2017
observed<-readRDS(here("data","AK_Effort_Trawl","Gasper_data","Unique_trips_Observed.RDS")) %>%
  rename_all(tolower) %>% 
  mutate(month=as.numeric(format(haul_date,"%m")),
         trip_target_code2=ifelse(akr_gear_code=="NPT","Not-P",trip_target_code),
           region = case_when(reporting_area_code == "610" ~ "E.APEN", 
                            reporting_area_code == "620" ~ "W.APEN",
                            reporting_area_code %in% c("630") ~ "NW.GOA",
                            reporting_area_code %in% c("640", "640 ") ~ "NE.GOA",
                            TRUE ~ "MISSING_LOCATION_INFO")) %>% 
  group_by(year,catcher_boat_adfg,pscnq_processing_sector,trip_target_code, month,trip_target_code2,akr_gear_code,region, reporting_area_code) %>%
  summarise(haul_days=length(unique(haul_date[!is.na(haul_date)])),
            vlength=ves_cfec_length[1],
            gf_catch=sum(official_total_catch,na.rm=TRUE)) %>%
  filter(case_when(year == 1996 ~ !akr_gear_code == "NPT",
                   TRUE ~ trip_target_code == "P"), !reporting_area_code == "650") %>% #1996 trips are classified slightly differently, tyhere is no trip target code so remove NPT for 1996 only. #only 1 entry from 650 dont know if it is SSEAK or NSEAK
  ungroup() %>%
  mutate(
    # length_category = case_when(vlength < 125 & year < 2013 ~ "Small",
    #                                  vlength > 124 & year < 2013 ~ "Large",
    #                                  vlength > 59 & year > 2012 ~ "Large",
    #                                  vlength < 60 & year > 2012 ~ "Small",
    #                                  TRUE ~ as.character(vlength)),
          # season = case_when(month %in% c(1,2,3,11,12) ~ "winter",
          #                       month %in% c(4, 5) ~ "spring",
          #                       month %in% c(6,7) ~ "summer",
          #                       month %in% c(8,9,10) ~ "fall",
          #                       TRUE~"SEASON_NA"),
         year = as.numeric(year)) %>%  #add season for plotting purposes- season structure = what ole uses in 2020 paper
  group_by(year,pscnq_processing_sector,trip_target_code, month,region
           #,length_category, #can add this in later if fish ticket data for obs and unobs has v length?
           ) %>%
  summarise(observed_boat_days = sum(haul_days),
             obs_catch_MT=sum(gf_catch)) #THIS IS IN MT--CONVERT TO LBS OR FISH TICKET CONVERT TO MT

#all, observed and unobserved trips ***shoreside only
fish_ticket<-readRDS(here("data","AK_Effort_Trawl","Gasper_data","fishtix_snippet.RDS")) %>% 
  mutate(fishing_start_date=as.Date(ft_date_fishing_began,format="%Y%m%d"), # the dates are wonky - convert to acutal dates
         date_landed=as.Date(ft_date_landed,format="%Y%m%d"),
         region = case_when(reporting_area_code == "610" ~ "E.APEN", 
                            reporting_area_code == "620" ~ "W.APEN",
                            reporting_area_code %in% c("630") ~ "NW.GOA",
                            reporting_area_code %in% c("640", "640 ") ~ "NE.GOA",
                            TRUE ~ "MISSING_LOCATION_INFO")) %>% 
  group_by(vessel_id,adfg_number,date_landed,year,region) %>% # group by vessel-date_landed-year,reporting_area_code
  summarise(ft_observers_onboard=ft_observers_onboard[1], # this observers_onboard field seems pretty useless.
              observed=observed[1], #  this flag for whether a vessel is observed seems somewhat more useful. Maybe you'll find it helpful?
              fishing_start_date=min(fishing_start_date), # for a particular landed date there may be multiple start dates. Pick the earliest one.
              pounds=sum(pounds[earnings>0],na.rm=TRUE)) %>%  # summarise pounds of only those species for which they got paid
  data.frame %>% 
  rowwise() %>% 
  mutate(fish_days=as.vector(difftime(date_landed,fishing_start_date,units="days")), # determine number of days between start and landed. Note that if landed the same day it will be a zero. Maybe you want to add 1? [line below]
         fish_days = case_when(fishing_start_date==date_landed ~ 1,
                               TRUE ~ fish_days),
         month=as.numeric(format(date_landed,"%m")),
         year=as.numeric(format(date_landed,"%Y")),
         # season = case_when(month %in% c(1,2,3,11,12) ~ "winter",
         #                        month %in% c(4, 5) ~ "spring",
         #                        month %in% c(6,7) ~ "summer",
         #                        month %in% c(8,9,10) ~ "fall",
         #                        TRUE~"SEASON_NA")
         ) %>% # extract a month and year field
 group_by(year,month,region) %>% #  group by vessel-date_landed-year,reporting_area_code
  summarise(total_effort = sum(fish_days),
             all_catch_MT=sum(pounds*0.000453592)) %>% #convert to MT, 0.000453592 MT = 1 LB 
filter(!region == "MISSING_LOCATION_INFO") 

shoreside_observed <- observed %>%
  filter(pscnq_processing_sector == "S")

#merge data and get the fraction of obs and unobserved. 
join_data_shoreside <- left_join(fish_ticket, shoreside_observed, by=c("year", "month", "region")) %>%
  dplyr::mutate(observed_boat_days = as.numeric(replace_na(observed_boat_days,0)),
                total_effort = replace_na(total_effort,0),
                unobs_boat_days = total_effort - observed_boat_days)  %>%
  filter(!region == "MISSING_LOCATION_INFO")  %>%
  ungroup() %>%
  mutate(observed_boat_days = case_when(year<1996 ~ unobs_boat_days*0.3, #0.3 based on oberver coverage chart when we dont have data for 1991-1996.
                                         TRUE ~observed_boat_days),
          unobs_boat_days = case_when(year<1996 ~ unobs_boat_days-observed_boat_days, TRUE ~ unobs_boat_days),
          frac_boat_days_obs = observed_boat_days/total_effort,
          frac_boat_days_obs = replace_na(frac_boat_days_obs, 0))
```

\n Stacked bar for observed vs unobserved effort all years and months- x axis is year.month

```{r stacked bar plots new, echo=FALSE, fig.height=5, fig.width=9}
temp.year <- as.character(c(1991:2017))
temp.month <- as.character(c(01:12))
temp.region <-  as.character(c("E.APEN", "W.APEN",  "NW.GOA", "NE.GOA"))
category <-  as.character(c("unobs_boat_days", "observed_boat_days"))
year.month<- expand.grid(temp.year, temp.month,category,temp.region) %>%  #fully factorial match between all categories 
rename(category="Var3", region="Var4")
year.month$Var2<- str_pad(year.month$Var2, width= 2, pad = "0", side="left") #add a zero before month 1-9 so that I can keep months in order
year.month <- unite(year.month, "year.month", c("Var1", "Var2"), sep = ".", remove= TRUE) #join month and year in its own dataframe 

#general plot of all effort. 
join_data_shoreside %>%  
  ungroup() %>%
  select(year,month,region,pscnq_processing_sector,unobs_boat_days,observed_boat_days,frac_boat_days_obs) %>%   
  mutate(month = str_pad(month, "left", pad=0, width = 2)) %>%
  unite("year.month",c("year","month"), sep = ".") %>%
  gather(4:5, key = "category", value = "boat_days") %>%
  merge(year.month, all = TRUE) %>% # add in labels
  mutate(boat_days= replace_na(boat_days,0)) %>%
  filter(!boat_days<0) %>%
    ggplot() +
      geom_bar(aes(x=year.month,y=boat_days, fill = category), position="stack", stat="identity",alpha = 0.8) +
      facet_wrap(~region, scales = "free", ncol=1) +
      ggtitle("Boat Days (Shoreside Pollock)") + # no small CP's
      theme_classic() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1,size=7))  +
      scale_x_discrete(breaks = every_nth(n = 12))
 
```

\n Stacked bar plots summarized by season
```{r plot seperately by seasons }
#make specific x-axes
temp.region <-  as.character(c("E.APEN", "W.APEN",  "NW.GOA", "NE.GOA"))
category <-  as.character(c("unobs_boat_days", "observed_boat_days"))

temp<- expand.grid(temp.year, category,temp.region) %>%  #fully factorial match between all categories 
  rename(year="Var1",category="Var2", region="Var3") #%>%
  #mutate(Var2= str_pad(Var2, width= 2, pad = "0", side="left")) #%>%
#  unite("year.month", c("Var1", "Var2"), sep = ".", remove= TRUE) #join month and year in its own dataframe 
```

```{r SPRING new}
#SPRING
join_data_shoreside %>%  
  ungroup() %>%
  filter(month %in% c(4,5)) %>%
  select(1:3,8,10) %>%
  gather(4:5, key = "category", value = "boat_days") %>%
  group_by(year,region,category) %>%
  summarise(boat_days=sum(boat_days))%>%
  ungroup() %>%
  merge(temp, all = TRUE) %>%
  mutate(boat_days= replace_na(boat_days,0)) %>%
   ggplot() +
    geom_bar(aes(x=year,y=boat_days, fill = category), position="stack", stat="identity",alpha = 0.8) +
    facet_grid(~region, scales = "free") +
    ggtitle("Spring Boat Days (Shoreside Pollock)") + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1,size=7))  +
  scale_x_discrete(breaks = every_nth(n = 5))

```

```{r SUMMER new}
#SUMMER
join_data_shoreside %>%  
  ungroup() %>%
  filter(month %in% c(6,7)) %>%
    select(1:3,8,10) %>%
  gather(4:5, key = "category", value = "boat_days") %>%
  group_by(year,region,category) %>%
  summarise(boat_days=sum(boat_days))%>%
  ungroup() %>%
  merge(temp, all = TRUE) %>%
  mutate(boat_days= replace_na(boat_days,0)) %>%
   ggplot() +
    geom_bar(aes(x=year,y=boat_days, fill = category), position="stack", stat="identity",alpha = 0.8) +
    facet_grid(~region, scales = "free") +
    ggtitle("Summer Boat Days (Shoreside Pollock)") + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1,size=7))  +
  scale_x_discrete(breaks = every_nth(n = 5))
 
```

```{r FALL new}
#FALL
join_data_shoreside %>%  
  ungroup() %>%
  filter(month %in% c(8,9,10)) %>%
  select(1:3,8,10) %>%
  gather(4:5, key = "category", value = "boat_days") %>%
  group_by(year,region,category) %>%
  summarise(boat_days=sum(boat_days))%>%
  ungroup() %>%
  merge(temp, all = TRUE) %>%
  mutate(boat_days= replace_na(boat_days,0)) %>%
   ggplot() +
    geom_bar(aes(x=year,y=boat_days, fill = category), position="stack", stat="identity",alpha = 0.8) +
    facet_grid(~region, scales = "free") +
    ggtitle("Fall Boat Days (Shoreside Pollock)") + 
    theme_classic() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1,size=7))  +
  scale_x_discrete(breaks = every_nth(n = 5))
```

```{r WINTER new}
 #WINTER
join_data_shoreside %>%  
  ungroup() %>%
  select(1:3,8,10) %>%
  filter(month %in% c(1,2,3,11,12)) %>%
  mutate(season.temp = case_when(month %in% c(11,12) ~ "winter.1",
                       month %in% c(1:3) ~ "winter.2",
                       TRUE ~ "NA")) %>%
  gather(4:5, key = "category", value = "boat_days") %>%
  group_by(year, season.temp,region,category) %>%
  summarise(boat_days=sum(boat_days)) %>%
  ungroup() %>%
  mutate(year=as.numeric(year), year.new = case_when(season.temp == "winter.2" ~ year -1,
                              TRUE ~ year)) %>%
  group_by(year.new, region, category) %>%
  summarise(boat_days=sum(boat_days)) %>%
  dplyr::rename(year=year.new) %>%
  merge(temp) %>%
  mutate(boat_days= replace_na(boat_days,0)) %>%
  ggplot() +
    geom_bar(aes(x=year,y=boat_days, fill = category), position="stack", stat="identity",alpha = 0.8) +
    facet_grid(~region, scales = "free") +
    ggtitle("Winter Boat Days ( Shoreside Pollock)") + # no small CP's
    theme_classic() +
    theme(axis.text.x = element_text(angle = 90, hjust = 1,size=7)) 

 
```

\n How long do trips tend to be?  
```{r trip length data}
all_data_trip_length<-readRDS(here("data","AK_Effort_Trawl","Gasper_data","fishtix_snippet.RDS")) %>% 
  mutate(fishing_start_date=as.Date(ft_date_fishing_began,format="%Y%m%d"), # the dates are wonky - convert to acutal dates
         date_landed=as.Date(ft_date_landed,format="%Y%m%d"),
         region = case_when(reporting_area_code == "610" ~ "E.APEN", 
                            reporting_area_code == "620" ~ "W.APEN",
                            reporting_area_code %in% c("630") ~ "NW.GOA",
                            reporting_area_code %in% c("640", "640 ") ~ "NE.GOA",
                            TRUE ~ "MISSING_LOCATION_INFO")) %>% 
  group_by(vessel_id,adfg_number,date_landed,year,region) %>% #  group by vessel-date_landed-year,reporting_area_code
  summarise(ft_observers_onboard=ft_observers_onboard[1], # this observers_onboard field seems pretty useless.
              observed=observed[1], #  this flag for whether a vessel is observed seems somewhat more useful. Maybe you'll find it helpful?
              fishing_start_date=min(fishing_start_date), # for a particular landed date there may be multiple start dates. Pick the earliest one.
              pounds=sum(pounds[earnings>0],na.rm=TRUE)) %>%  # summarise pounds of only those species for which they got paid
  data.frame %>% 
  rowwise() %>% 
  mutate(fish_days=as.vector(difftime(date_landed,fishing_start_date,units="days")), # determine number of days between start and landed. Note that if landed the same day it will be a zero. Maybe you want to add 1? [line below]
         fish_days = case_when(fishing_start_date==date_landed ~ 1,
                               TRUE ~ fish_days),
         month=as.numeric(format(date_landed,"%m")),
         year=as.numeric(format(date_landed,"%Y")),
         season = case_when(month %in% c(1,2,3,11,12) ~ "winter",
                                month %in% c(4, 5) ~ "spring",
                                month %in% c(6,7) ~ "summer",
                                month %in% c(8,9,10) ~ "fall",
                                TRUE~"SEASON_NA")) %>%
  group_by(year, region) %>%
  summarise(mean_trip_length = mean(fish_days),
            sd_fish_days=sd(fish_days)) %>%
  filter(!region == "MISSING_LOCATION_INFO")

all_data_trip_length %>% 
  ungroup() %>%
 # mutate(month = str_pad(month, "left", pad=0, width = 2)) %>%
  #unite("year.month",c("year","month"), sep = ".") %>%
ggplot(aes(x=year,y=mean_trip_length)) +
   geom_line()+ 
   geom_point(alpha = 0.8) +
  #geom_errorbar(aes(ymin=mean_trip_length-sd_fish_days, ymax=mean_trip_length+sd_fish_days))+
   facet_wrap(~region, scales = "free", ncol=1) +
   ggtitle("Spring Boat Days (Shoreside Pollock)") + 
   theme_classic() +
   theme(axis.text.x = element_text(angle = 90#, hjust = 1,size=7
                                    ))  
```

Fraction of Boat Days with an observer onboard.
Black = no effort, white = no observed effort

```{r heatmap obs frac}
spread.total <- join_data_shoreside %>%
  ungroup() %>%
  select(c(1:3,11)) %>%
  mutate(month = str_pad(month, "left", pad=0, width = 2)) %>%
  unite("year.month",1:2, remove=TRUE, sep = ".") %>%
  # group_by(year.month, region) %>%
  # summarise(total_effort = sum(total_effort),
  #           observed_boat_days = sum(observed_boat_days)) %>%  
  merge(year.month %>% select(-c(category)) %>% group_by(year.month) %>% count(region) %>%select(-c(n)), all = TRUE) %>%
 # mutate(obs_frac= observed_boat_days/total_effort) %>%
  filter(!frac_boat_days_obs >1) %>% #handful of errors as JW about. 
  dplyr::select(year.month,region, frac_boat_days_obs) %>%
  tidyr::spread(year.month, frac_boat_days_obs) 

#set up matrix for heatmap 
#spread.total<- spread.total[1:5,] 
region<- spread.total$region
spread.total<- spread.total[,2:237]
row.names(spread.total) <- region
matrix= as.matrix(spread.total)
#matrix[matrix < 0.1] <- NA #turn 0 to NA so that it doesnt get messed up in log function

# FOR X AXIS LABELS
temp.lab = as.data.frame(temp.year) #create df with all years
temp <- as.data.frame(temp.lab[rep(1:nrow(temp.lab),1,each=12),]) #duplicate each row 12 times
temp$`temp.lab[rep(1:nrow(temp.lab), 1, each = 12), ]`[duplicated(temp$`temp.lab[rep(1:nrow(temp.lab), 1, each = 12), ]`)] <- " " #fill duplicates with a space
temp$label<-  temp$`temp.lab[rep(1:nrow(temp.lab), 1, each = 12), ]`  #change column name and remove old column

heatmap.2(matrix, Rowv=FALSE, Colv=FALSE, 
          xlab = "Year.Month",  
          scale="none", margins=c(4,6), 
          dendrogram="none", trace="none",
          col=col.br(20), #breaks=pairs.breaks,
          na.color = "black",
          key =TRUE,
          keysize = 1.5, #labRow = year_list,
          key.title = "log scale",
          adjCol = c(0.5,1.1),
          labCol = temp$label, #tell it to use dummy variable for X axis years
          labRow = region, #use dummy variable for Y axis regions so they can be north to south 
          cexCol = 0.6,
          sepcolor = "black",
          main = "Fraction of Observed Boat Days", 
          density.info = ("none"),
          #( "bottom.margin", "left.margin", "top.margin", "right.margin" )
          key.par=list(mar=c(2,1, 1,0.1)))

```

\n If there is an observer on board- how much of the catch are they surveying? 10% unless a vessel delivers to a tender, than 0%--- Below is a summary of fraction of total Chinook sampled per trip, given there is an observer onboard. Any fraction less than 10% means that some trips during that month were delivered to a tender.

\n This data comes from "Chinooky_rates_merged.RDS", the tender delivery difference only started after ~2010 so these data are appropriate to use, earlier years are 10%

```{r sample frac 2}
chinooky_rates<- readRDS(here("data","AK_Effort_Trawl","Gasper_data","Chinooky_Rates_merged.RDS")) %>%
filter(fmp_sub_area_code == "GOA", target =="pollock") %>%
        mutate(region = case_when(reporting_area_code == "610" ~ "E.APEN",
                            reporting_area_code == "620" ~ "W.APEN",
                            reporting_area_code %in% c("630") ~ "NW.GOA",
                            reporting_area_code %in% c("640", "640 ") ~ "NE.GOA",
                            reporting_area_code == "649" & adfg_stat_area_code %in% c("466032","466033","466003") ~ "NE.GOA",
                            reporting_area_code == "649" & adfg_stat_area_code %in% c("476006","486001","476035","476005","476007","476003","476004","476034", "475933","476031","476008","485932") ~ "NW.GOA", #there are about 600 records from NMFS AREA 649, which is the most northern part of PWS stat area- going to divide these based on in the 630/640 line continued into PWS and assign via ADFG #'s
                            TRUE ~ "MISSING_LOCATION_INFO")) %>% #there are 14 of these because they are in area 649 but have missing ADFG #'s, all from a few days  in 2003
        separate(trip_target_date, into = c("year", "month", "day"), sep = "-")  
   
    temp.assign <- chinooky_rates %>%
            filter(region == "MISSING_LOCATION_INFO") %>%
            mutate(temp.numb = rep(1:2, times =7)) %>%
            mutate(region = case_when(region == "MISSING_LOCATION_INFO" & temp.numb == 1 ~ "NE.GOA",
                        region == "MISSING_LOCATION_INFO" & temp.numb == 2 ~ "NW.GOA",
                        TRUE ~ region)) %>%
            select(-c(temp.numb))
    chinooky_rates <- chinooky_rates %>%
      filter(!region == "MISSING_LOCATION_INFO") %>%
      rbind(temp.assign) 

#   How often do tender deliveries occur? 
frac_2<- chinooky_rates %>%
         filter(observers_onboard>0) %>%
         mutate(tender_vessel_present_absent=case_when(is.na(tender_vessel_adfg_number) ~ "ten_percent_obs",
                                                      TRUE ~"No_obs")) %>%
        group_by(year,month,region) %>%
        count(tender_vessel_present_absent) %>%
        spread(tender_vessel_present_absent,n) %>%
        ungroup() %>%     
        mutate(No_obs=replace_na(No_obs,0),frac_boat_days_obs_vs_tender = ten_percent_obs/(ten_percent_obs+No_obs),
                    overall_sample_frac = frac_boat_days_obs_vs_tender *.10, month = str_pad(month, "left", pad=0, width = 2) ) %>%
        unite("year.month",c("year","month"), sep = ".") 
   
  ggplot(frac_2) +
     geom_point(aes(x=year.month, y = overall_sample_frac)) +
     facet_wrap(~region) +
   theme_classic()+
   theme(axis.text.x = element_text(angle = 90, hjust = 1,size=7))  +
   scale_x_discrete(breaks = every_nth(n = 20))
      
```

\n Does effort generally line up with RMIS recoveries? [using RMIS recoveries from all stocks]
```{r rmis prep}

rmis<-readRDS(here("data","rmis_allstocks_May2020.RDS"))%>%
  filter(fishery == 81) #filter to just GOA trawl data 

pollock_rmis <- rmis %>%
   group_by(region, rec_year) %>%
   count(rec_month) %>%
  unite("year.month",c("rec_year","rec_month"), sep = ".")  %>%
  rename(recoveries =n)

```

\n Plot observed effort and recoveries- looks ok there aren't too many lone pink bars (indicates recovery occured but no data for the observed effort)

```{r rmis plot}
 join_data_shoreside %>%  
  ungroup() %>%
  select(year,month,region,observed_boat_days) %>%   
  mutate(month = str_pad(month, "left", pad=0, width = 2) ) %>%
  unite("year.month",c("year","month"), sep = ".") %>% 
  full_join(pollock_rmis, by= c("year.month","region")) %>%
  gather(3:4, key = "category", value = "number") %>%
  mutate(number= replace_na(number,0), category=factor(category, levels=c("recoveries", "observed_boat_days"))) %>%
  filter(!is.na(region), !region %in% c("NE.GOA NW.GOA W.APEN E.APEN BER", "NSEAK")) %>%
    ggplot() +
      geom_bar(aes(x=year.month,y=number, fill = category),position="stack", stat="identity",alpha = 0.8) +
      facet_wrap(~region, scales="free_y",ncol = 1) +
      ggtitle("Compare Observed Data and RMIS Recoveries") +  
      theme_classic() +
      theme(axis.text.x = element_text(angle = 90, hjust = 1,size=7))  +
      scale_x_discrete(breaks = every_nth(n = 20))
```
\n Heatmap plot of total sample fraction

```{r heatmap total frac}
spread.total.frac <- join_data_shoreside %>%
  ungroup() %>%
  select(c(year,month,region,frac_boat_days_obs)) %>%
  mutate(month = str_pad(month, "left", pad=0, width = 2)) %>%
  unite("year.month",1:2, sep = ".") %>%
  left_join(frac_2 %>% select(c(year.month, region,overall_sample_frac)), by=c("year.month", "region")) %>%
  mutate(overall_sample_frac=case_when(frac_boat_days_obs>0 & is.na(overall_sample_frac) ~ 0.1, 
                                       is.na(overall_sample_frac) ~ 0,
                                       TRUE~ overall_sample_frac), 
         total_fraction_effort_surveyed_for_chinook= overall_sample_frac*frac_boat_days_obs) %>%
  merge(year.month %>% select(-c(category)) %>% group_by(year.month) %>% count(region) %>%select(-c(n)), all = TRUE) %>%
  dplyr::select(year.month,region, total_fraction_effort_surveyed_for_chinook) %>%
  tidyr::spread(year.month, total_fraction_effort_surveyed_for_chinook) 

#set up matrix for heatmap 
#spread.total<- spread.total[1:5,] 
region<- spread.total.frac$region
spread.total.frac<- spread.total.frac[,2:325]
row.names(spread.total.frac) <- region
matrix= as.matrix(spread.total.frac)
#matrix[matrix < 0.1] <- NA #turn 0 to NA so that it doesnt get messed up in log function

# FOR X AXIS LABELS
temp.lab = as.data.frame(temp.year) #create df with all years
temp <- as.data.frame(temp.lab[rep(1:nrow(temp.lab),1,each=12),]) #duplicate each row 12 times
temp$`temp.lab[rep(1:nrow(temp.lab), 1, each = 12), ]`[duplicated(temp$`temp.lab[rep(1:nrow(temp.lab), 1, each = 12), ]`)] <- " " #fill duplicates with a space
temp$label<-  temp$`temp.lab[rep(1:nrow(temp.lab), 1, each = 12), ]`  #change column name and remove old column

heatmap.2(matrix, Rowv=FALSE, Colv=FALSE, 
          xlab = "Year.Month",  
          scale="none", margins=c(4,6), 
          dendrogram="none", trace="none",
          col=col.br(20), #breaks=pairs.breaks,
          na.color = "black",
          key =TRUE,
          keysize = 1.5, #labRow = year_list,
          key.title = "log scale",
          adjCol = c(0.5,1.1),
          labCol = temp$label, #tell it to use dummy variable for X axis years
          labRow = region, #use dummy variable for Y axis regions so they can be north to south 
          cexCol = 0.6,
          sepcolor = "black",
          main = "Total Fraction of Effort Surveyed\nfor CWT Chinook in Shoreside", 
          density.info = ("none"),
          #( "bottom.margin", "left.margin", "top.margin", "right.margin" )
          key.par=list(mar=c(2,1, 1,0.1)))

```
 
```{r, eval = FALSE, echo=FALSE}

CP_observed <- observed %>%
  filter(pscnq_processing_sector %in% c("M", "CP"))

observed<-readRDS(here("data","AK_Effort_Trawl","Gasper_data","Unique_trips_Observed.RDS")) %>%
  rename_all(tolower) %>% 
  mutate(month=as.numeric(format(haul_date,"%m")),
         trip_target_code2=ifelse(akr_gear_code=="NPT","Not-P",trip_target_code),
           region = case_when(reporting_area_code == "610" ~ "E.APEN", 
                            reporting_area_code == "620" ~ "W.APEN",
                            reporting_area_code %in% c("630") ~ "NW.GOA",
                            reporting_area_code %in% c("640", "640 ") ~ "NE.GOA",
                            TRUE ~ "MISSING_LOCATION_INFO")) %>% 
  group_by(year,catcher_boat_adfg,pscnq_processing_sector,trip_target_code, month,trip_target_code2,akr_gear_code,region, reporting_area_code) %>%
  summarise(haul_days=length(unique(haul_date[!is.na(haul_date)])),
            vlength=ves_cfec_length[1],
            gf_catch=sum(official_total_catch,na.rm=TRUE)) %>%
  filter(case_when(year == 1996 ~ !akr_gear_code == "NPT",
                   TRUE ~ trip_target_code == "P"), !reporting_area_code == "650")

```

```{r save data }
#shoreside only 
#save effort data
 final_shoreside<- join_data_shoreside %>%
  ungroup() %>%
  select(year,month,region,unobs_boat_days,observed_boat_days,total_effort)  

saveRDS(final_shoreside, "data/AK_Effort_Trawl/SHoreside_Pollock_Effort_Summarized.RDS")

#if there is an observer onboard** [so this is filtered for just observed trips] they sample 1/10 1996 and lter, unless specific tender deliver stuff happens 
#save sample fraction data 
 total.frac <- join_data_shoreside %>%
  ungroup() %>%
  select(c(year,month,region,frac_boat_days_obs)) %>%
  mutate(month = str_pad(month, "left", pad=0, width = 2)) %>%
  unite("year.month",1:2, sep = ".") %>%
  left_join(frac_2 %>% select(c(year.month, region,overall_sample_frac)), by=c("year.month", "region")) %>%
  filter(!frac_boat_days_obs == 0 ) %>%
  mutate(sample_fraction=case_when(is.na(overall_sample_frac) ~ 0.1, 
                                TRUE ~ overall_sample_frac)) %>%
  select(-frac_boat_days_obs, -overall_sample_frac)
  
saveRDS(total.frac, "data/AK_Effort_Trawl/Sample_Fraction_Pollock_Summarized.RDS")


   



```

